Weighted Averaging of Multiple Non-collocated Inertial Measurement Units

### 2024-08-28 paper/main.tex
Proof read and small fixes to the paper.

% additional packages
\usepackage{bm}
\usepackage{mathtools}

@@

\begin{abstract}
Multi-IMU fusion has been an effective way of improving MEMS IMU accuracy. In this article, we discuss the averaging technique with multiple IMUs with significant separation. We also present methods of selecting weighting terms such that we either minimize the uncertainty in fused sensor output or arbitrarily placing the reference frame of the combined virtual IMU (VIMU). We tested our method with simulation and evaluated on real dataset. The result shows that the averaging technique works for IMUs with separation and performance gain is observed.
\end{abstract}

@@

Among many sensor configuration with IMU, monocular camera plus IMU is one of the minimal configuration that is well-studied \cite{qin2018_vins-mono, 10616216}. Also IMU has been used heavily as a complimentary sensor for GPS in GNSS-denied environment \cite{1008998, 8987949}. Based on these minimal configurations, various extension has been made. UMS-VINS \cite{jiang2023_UMS-VINS} provides a unified framework for using monocular and stereo cameras in visual-inertial odometry. Also the MSCKF framework mentioned above \cite{Anastasios2007_MSCKF} also works for multiple cameras with a IMU. There are also works that aim to provide a complete framework to incorporate multiple different sensors in one unified system \cite{10587194}.

@@

One category of approach is to perform fusion of multiple IMU before using the measurement in inertial navigation system (INS). For example, \cite{jafari2014_PEM} uses Prediction Error Minimization method to model noise of IMUs and use Kalman filter to estimate the optimal process signal. \cite{xue2023} proposed to use Kalman filter to estimate the angular velocity measured by multiple non-orthogonal gyroscopes. Instead of using Kalman filter, an easier approach would be to directly average multiple gyroscopes after aligned under a common virtual IMU frame as discussed in \cite{waegli2008, patel2022_multi-imu, Colomina2004REDUNDANTIF}. Since VIMU frame is not the same for redundant accelerometer, thus for acceleration reading, due to the effect of additional acceleration from rotation, direct averaging was not considered previous unless IMUs are closed placed and separation in between are ignored.

@@

Typically state estimation problem with single IMU involves tracking the orientation, linear velocity and linear acceleration of the vehicle (commonly defined at the IMU). With additional IMUs, in addition to tracking robot state, additional state variables like linear acceleration, angular velocity and angular acceleration are added. This allows using the full mechanization model of accelerometer as measurement update in Kalman filter as in \cite{Bancroft2011DataFA, Beaudoin2018_satelite}.

@@

Another approach is to track individual IMU's state and associate them with geometric constraints \cite{waegli2008, Beaudoin2018_satelite}. Using multiple IMU and camera, \cite{Eckenhoff2021_MIMC-VINS} proposed to track all stacked states of base and auxillary IMUs using MSCKF \cite{Anastasios2007_MSCKF}. Assuming the calibrations are known, relative geometric constraints is imposed between states of different IMUs in the form of measurement update in addition to camera observation.

@@

Last common approach would be to have individual state estimator for each IMU system each perform local propagation and correction. Simple state fusion is performed with each estimator's estimate \cite{Bancroft2011DataFA, patel2022_multi-imu}.

@@

As presented, multiple approaches have being proposed. \ref{constraint} would be computationally expensive if number of IMU are large as it has many states all in one Kalman filter. The averaging approach in \ref{VIMU} is the simplest to implement and doesn't significantly alter the existing single IMU framework. Although averaging between accelerometers that are physically separated is not proposed due to the lever arm effect, in \ref{methodology} we present that it is still possible to average the accelerometers despite the separation.

@@

For any quantity $x$ We define $\bar{x} = x + b_x$ as the biased value and $\Tilde{x} = x + n_x$ as the noisy value. If it is both biased and noisy measurement, it is denoted as $y_x$. If a quantity is in the form like $x(t)$ then it is a time-varying quantity, otherwise it is time-invariant.

@@

For rotation matrix or transformation matrix, we have $\textbf{C}_{vi}$ or $\textbf{T}_{vi}$ for transforming point $\textbf{r}_i^{pi}$ expressed in inertial frame $\bm{\mathcal{F}}_i$ into vehicle frame $\bm{\mathcal{F}}_v$ as $\textbf{r}_v^{pi}$.
$$
\textbf{r}_v^{pi} = \textbf{T}_{vi} \textbf{r}_i^{pi}
$$

@@

The state propagation using IMU measurement follows the rigid-body kinematics in world fixed frame.
$$
\begin{equation}
\begin{align}
    \dot{\textbf{C}}(t) &= \textbf{C}(t) \bm{\omega}(t)^\wedge \\
    \dot{\textbf{v}}(t) &= \textbf{C}(t) \textbf{a}(t) - \textbf{g} \\
    \dot{\textbf{p}}(t) &= \textbf{v}(t)
\end{align}
\end{equation}
$$

@@

Where $\textbf{b}_\omega$ is the bias of gyroscope measurement due to random walk and $\textbf{n}_\omega$ is the additive white noise on the measurement. Similarly we have the following measurement model for IMU at vehicle frame. Here $\textbf{b}_a$ is the bias of accelerometer and $\textbf{n}_a$ is the noise on the acceleration measurement.

@@

@@

For IMU not at vehicle frame, measurement are in each IMU's local frame and we also have additional terms in acceleration measurement due to lever arm effect.

\begin{equation}
    \textbf{y}_{\omega,i}(t) = \textbf{C}_{i}^{-1} \bm{\omega}(t) + \textbf{b}_{\omega, i}(t) + \textbf{n}_{\omega,i}(t)
\end{equation}
\begin{equation}
\begin{split}
    \textbf{y}_{a,i}(t) &= \textbf{C}_i^{-1} \left[ \textbf{a}(t) + \bm{\omega}(t)^\wedge (\bm{\omega}(t)^\wedge \textbf{r}_i) + \bm{\alpha}(t)^\wedge \textbf{r}_i \right] \\
    &+ \textbf{b}_{a,i}(t) + \textbf{n}_{a,i}(t)
\end{split}
\end{equation}

@@

First we align gyroscope measurement with vehicle frame.
\begin{equation}
    \textbf{C}_{i} \textbf{y}_{\omega,i}(t) = \bm{\omega}(t) + \textbf{C}_{i} \textbf{b}_{\omega, i}(t) + \textbf{C}_{i} \textbf{n}_{\omega,i}(t)
\end{equation}

@@

Align the all accelerometer measurement to vehicle frame.
\begin{equation}\label{eqn_accel}
\begin{split}
    \textbf{C}_{i} \textbf{y}_{a,i}(t) &= \textbf{a}(t) + \bm{\omega}(t)^\wedge (\bm{\omega}(t)^\wedge \textbf{r}_i) + \bm{\alpha}(t)^\wedge \textbf{r}_i \\
    &+ \textbf{C}_{i} \textbf{b}_{a,i}(t) + \textbf{C}_{i} \textbf{n}_{a,i}(t)
\end{split}
\end{equation}

Summing equation \ref{eqn_accel} of all accelerometers we have,
\begin{equation}
\begin{split}
    \bar{\textbf{y}}_a(t) &= \sum_i{a_i \textbf{C}_{i} \textbf{y}_a(t)} \\
    &= \textbf{a}(t) + \sum_i{a_i \textbf{C}_{i} \textbf{b}_{a,i}(t)} + \sum_i{a_i \textbf{C}_{i} \textbf{n}_{a,i}(t)} \\
    &+ \sum_i{a_i \bm{\omega}(t)^\wedge (\bm{\omega}(t)^\wedge \textbf{r}_i)} + \sum_i{a_i \bm{\alpha}(t)^\wedge \textbf{r}_i} \\
\end{split}
\end{equation}

@@

If $\sum_i{a_i \textbf{r}_i} = \textbf{0}$ then we eliminate the last two terms from lever arm effect. This gives,
\begin{equation}
    \bar{\textbf{y}}_a(t) = \textbf{a}(t) + \bar{\textbf{b}}_a(t) + \bar{\textbf{n}}_a(t)
\end{equation}

@@

As we can see that, if we choose the vehicle frame and weights such that $\sum_i{a_i \textbf{r}_i} = \textbf{0}$, then we can ignore the separation of all IMUs and fused the accelerometer measurement through averaging.

Since all biases are slow varying, the combined biases are also slow varying. So instead of tracking individual biases, only the combined biases are needed. When receiving new set of IMU measurement, we estimate the angular velocity and linear acceleration with the update equations below.
\begin{equation}
\begin{split}
    \hat{\bm{\omega}}(t) = \bar{\textbf{y}}_\omega(t) - \hat{\bar{\textbf{b}}}_\omega(t) \\
    \hat{\textbf{a}}(t)  = \bar{\textbf{y}}_a(t) - \hat{\bar{\textbf{b}}}_a(t)
\end{split}
\end{equation}

@@

If the selection of vehicle frame is flexible and the goal is to minimize the noise in combined measurement or slowest bias drift. Below we take the bias drift as an example but similar idea can be applied to the output noise.

Bias drifts of IMU are modeled as random walk with the bias drift rate sampled from a white noise Gaussian.
\begin{equation}
\begin{split}
    \dot{\textbf{b}}_\omega(t) &= \textbf{w}_\omega(t) \\
    \dot{\textbf{b}}_a(t) &= \textbf{w}_a(t)
\end{split}
\end{equation}

@@

Since most IMU have isotropic noise for all three axes, the combined variance is simply,
\begin{equation}
    \bar{\sigma}^2 = \sum_i{a_i^2 \sigma_i^2}
\end{equation}

\subsection{Selecting Weight for Frame of Reference Placement}

Given that $\sum_i{a_i \textbf{r}_i} = \textbf{0}$, if there are only two IMUs, we can only place the vehicle frame on the line through the center of the two IMUs. If there are three IMUs, we can only place the vehicle frame on the plane form by the three IMUs. If there are four non-coplanar IMUs then in theory we can place the vehicle frame anywhere. However, placing the vehicle frame outside of the convex hull of the IMUs will increase the output noise which is usually not desired.

When we have a reachable vehicle frame in mind, we need to solve for the weight. Assuming we have arbitrarily many IMUs, then the problem can be formulated as below. In this section we will be using $w$ as weight instead of $a$ for convenience.
\begin{equation}
\begin{split}
    \min_{w_i}{\frac{1}{2} \sum{w_i^2 \sigma_i^2}} \quad \text{s.t. }
    \begin{cases}
      \sum_i{w_i \textbf{r}_i} &= \textbf{0} \\
      \sum_i{w_i} &= 1
    \end{cases}
\end{split}
\end{equation}

$\textbf{r}_i$ is the relative position of the IMUs to the selected vehicle frame. Here we used the objective to reduce noise as shown in section \ref{solve_weight_by_noise}.

Since it is a quadratic programming problem, it is possible to find a close-form solution. The Lagrangian of the problem is then $\mathcal{L} = \frac{1}{2}\textbf{w}^T \bm{\Sigma} \textbf{w} + \bm{\lambda}^T \textbf{R}\textbf{w} + \beta \left( 1 - \textbf{1}^T \textbf{w} \right)$ = 0 where $\bm{\Sigma} = \text{diag}(\sigma_1^2, ..., \sigma_n^2)$, $\textbf{R} = \left[\begin{matrix} \textbf{r}_0 & ... & \textbf{r}_n\end{matrix}\right]$ and $\textbf{1}^T = \left[\begin{matrix}1 & ... & 1\end{matrix}\right]$. This gives the following KKT conditions.
\begin{equation}
\begin{cases}
  \bm{\Sigma}\textbf{w} + \textbf{R}^T \bm{\lambda} - \beta \textbf{1} &= \textbf{0} \\
  \textbf{R} \textbf{w} &= \textbf{0} \\
  \textbf{1}^T \textbf{w} &= 1
\end{cases}
\end{equation}

@@

Lastly since $\sum_i{w_i} = 1$ we need to normalize the intermediate solution to get the final weight.
\begin{equation}
    \textbf{w}^* = \frac{\hat{\textbf{w}}}{\textbf{1}^T \hat{\textbf{w}}}
\end{equation}

\section{Experiments}

\subsection{Simulation}

Discuss how we generated the simulated trajectory and sensor measurement. Show the evaluation result (what metrics). Discuss about the noise reduction against theory.

@@

Which dataset we used (hard to find dataset containing multiple IMUs). What VINS implementation we selected. How we generate the averaged measurement (). The result compared to using single IMU. Does it show benefit of averaging?

>>>>

% additional packages
\usepackage{bm}
\usepackage{mathtools}
\usepackage{booktabs}
\usepackage{subcaption}

@@

\begin{abstract}
Multi-IMU fusion has been an effective way of improving MEMS IMU accuracy. In this article, we discuss the averaging technique for multiple IMUs with significant separation. We also present methods of selecting weighting terms such that we either minimize the uncertainty in fused sensor output or arbitrarily placing the reference frame of the combined virtual IMU (VIMU). We tested our method with simulation and evaluated on real dataset. The result shows that the averaging technique works for IMUs with separation and performance gain is observed.
\end{abstract}

@@

Among many sensor configuration with IMU, monocular camera plus IMU is one of the minimal configuration that is well-studied \cite{qin2018_vins-mono, 10616216}. Also IMU has been used heavily as a complimentary sensor for GPS in GNSS-denied environment \cite{1008998, 8987949}. Based on these minimal configurations, various extension has been made. UMS-VINS \cite{jiang2023_UMS-VINS} provides a unified framework for using monocular and stereo cameras in visual-inertial odometry. Also the MSCKF framework \cite{Anastasios2007_MSCKF} also works for multiple cameras with a IMU. There are also works that aim to provide a complete framework to incorporate multiple different sensors in one unified system \cite{10587194}.

@@

One category of approach is to perform fusion of multiple IMU before using the measurement in inertial navigation system (INS). For example, \cite{jafari2014_PEM} uses Prediction Error Minimization method to model noise of IMUs and use Kalman filter to estimate the optimal process signal. \cite{xue2023} proposed to use Kalman filter to estimate the angular velocity measured by multiple non-orthogonal gyroscopes. Instead of using Kalman filter, an easier approach would be to directly average multiple gyroscopes after aligned under a common virtual IMU frame as discussed in \cite{waegli2008, patel2022_multi-imu, Colomina2004REDUNDANTIF}.

However, averaging of multiple IMUs was not considered due to the additional acceleration terms from rotation (lever arm effect), unless they are closed place and the separation in between is ignored.

@@

Typically state estimation problem with single IMU involves tracking the orientation, linear velocity and linear acceleration of the vehicle (commonly defined at the IMU). With additional IMUs, additional state variables like linear acceleration, angular velocity and angular acceleration are added. This allows using the full mechanization model of accelerometer as measurement update in Kalman filter as in \cite{Bancroft2011DataFA, Beaudoin2018_satelite}.

@@

Another approach is to track individual IMU's state and associate them with geometric constraints \cite{waegli2008, Beaudoin2018_satelite}. Using multiple IMUs and cameras, \cite{Eckenhoff2021_MIMC-VINS} proposed to track all stacked states of base and auxillary IMUs using MSCKF \cite{Anastasios2007_MSCKF}. Assuming the calibrations are known, relative geometric constraints is imposed between states of different IMUs in the form of measurement update in addition to camera observations.

@@

Last common approach would be to have individual state estimator for each IMU system each performing local propagation and correction. Simple state fusion is performed with each estimator's estimate \cite{Bancroft2011DataFA, patel2022_multi-imu}.

@@

As presented, multiple approaches have being proposed. \ref{constraint} would be computationally expensive if number of IMU are large as it has many states all in one Kalman filter. The averaging approach in \ref{VIMU} is the simplest to implement and doesn't significantly alter the existing single IMU framework. Although averaging between accelerometers that are physically separated is not visited due to the lever arm effect, in \ref{methodology} we present that it is still possible to average the accelerometers despite the separation.

@@

For any quantity $x$ We define $\bar{x} = x + b_x$ as the biased value and $\Tilde{x} = x + n_x$ as the noisy value. If it is both biased and noisy measurement, it is denoted as $y_x$. If a quantity is in the form $x(t)$ then it is a time-varying quantity, otherwise it is time-invariant.

@@

For rotation matrix or transformation matrix, we have $\textbf{C}_{vi}$ or $\textbf{T}_{vi}$ for transforming point $\textbf{r}_i^{pi}$ expressed in inertial frame $\bm{\mathcal{F}}_i$ into vehicle frame $\bm{\mathcal{F}}_v$ as $\textbf{r}_v^{pv}$.
$$
\left[\begin{matrix} \textbf{r}_v^{pv} \\ 1 \end{matrix}\right]
    = \textbf{T}_{vi} \left[\begin{matrix} \textbf{r}_i^{pi} \\ 1 \end{matrix}\right]
\quad \text{or} \quad
\textbf{r}_v^{pv} = \textbf{C}_{vi} \textbf{r}_i^{pi} + \textbf{r}_v^{iv}
$$

@@

The state propagation using IMU measurement follows the rigid-body kinematics in world fixed frame.
\begin{equation}
\begin{align}
    \dot{\textbf{C}}(t) &= \textbf{C}(t) \bm{\omega}(t)^\wedge \\
    \dot{\textbf{v}}(t) &= \textbf{C}(t) \textbf{a}(t) - \textbf{g} \\
    \dot{\textbf{p}}(t) &= \textbf{v}(t)
\end{align}
\end{equation}

@@

Where $\textbf{b}_\omega$ is the bias of gyroscope measurement due to random walk and $\textbf{n}_\omega$ is the additive white noise of the measurement. Similarly we have the following measurement model for IMU at vehicle frame. Here $\textbf{b}_a$ is the bias of accelerometer and $\textbf{n}_a$ is the noise of the acceleration measurement.

@@

Bias drifts of IMU are modeled as random walk with the bias drift rate sampled from a Gaussian noise.
\begin{equation}
\begin{split}
    \dot{\textbf{b}}_\omega(t) &= \textbf{w}_\omega(t) \\
    \dot{\textbf{b}}_a(t) &= \textbf{w}_a(t)
\end{split}
\end{equation}

@@

For IMU not at vehicle frame, measurement are in each IMU's local frame and we also have additional terms in acceleration measurement due to lever arm effect.
\begin{equation}
    \textbf{y}_{\omega,i}(t) = \textbf{C}_{i}^{-1} \bm{\omega}(t) + \textbf{b}_{\omega, i}(t) + \textbf{n}_{\omega,i}(t)
\end{equation}
\begin{equation}
\begin{split}
    \textbf{y}_{a,i}(t) &= \textbf{C}_i^{-1} \left[ \textbf{a}(t) + \bm{\omega}(t)^\wedge (\bm{\omega}(t)^\wedge \textbf{r}_i) + \bm{\alpha}(t)^\wedge \textbf{r}_i \right] \\
    &+ \textbf{b}_{a,i}(t) + \textbf{n}_{a,i}(t)
\end{split}
\end{equation}

@@

First we align gyroscope measurements with vehicle frame.
\begin{equation}
    \textbf{C}_{i} \textbf{y}_{\omega,i}(t) = \bm{\omega}(t) + \textbf{C}_{i} \textbf{b}_{\omega, i}(t) + \textbf{C}_{i} \textbf{n}_{\omega,i}(t)
\end{equation}

@@

Align the all accelerometer measurements to vehicle frame.
\begin{equation}\label{eqn_accel}
\begin{split}
    \textbf{C}_{i} \textbf{y}_{a,i}(t) &= \textbf{a}(t) + \bm{\omega}(t)^\wedge (\bm{\omega}(t)^\wedge \textbf{r}_i) + \bm{\alpha}(t)^\wedge \textbf{r}_i \\
    &+ \textbf{C}_{i} \textbf{b}_{a,i}(t) + \textbf{C}_{i} \textbf{n}_{a,i}(t)
\end{split}
\end{equation}

Summing equation (\ref{eqn_accel}) of all accelerometer we have,
\begin{equation}
\begin{split}
    \bar{\textbf{y}}_a(t) &= \sum_i{a_i \textbf{C}_{i} \textbf{y}_a(t)} \\
    &= \textbf{a}(t) + \sum_i{a_i \textbf{C}_{i} \textbf{b}_{a,i}(t)} + \sum_i{a_i \textbf{C}_{i} \textbf{n}_{a,i}(t)} \\
    &+ \sum_i{a_i \bm{\omega}(t)^\wedge (\bm{\omega}(t)^\wedge \textbf{r}_i)} + \sum_i{a_i \bm{\alpha}(t)^\wedge \textbf{r}_i} \\
\end{split}
\end{equation}

@@

If $\sum_i{a_i \textbf{r}_i} = \textbf{0}$ then we can eliminate the last two terms from lever arm effect. This gives,
\begin{equation}
    \bar{\textbf{y}}_a(t) = \textbf{a}(t) + \bar{\textbf{b}}_a(t) + \bar{\textbf{n}}_a(t)
\end{equation}

@@

As we can see, if we choose the vehicle frame and weights of accelerometers such that $\sum_i{a_i \textbf{r}_i} = \textbf{0}$, then we can ignore the separation of all IMUs and fused the accelerometer measurement through averaging.

\textbf{Note that we can choose different weights for averaging gyroscope and accelerometers.}

Since all biases are slow-varying, the combined biases are also slow-varying. So instead of tracking individual biases, only the combined biases are needed. When receiving new set of IMU measurement, we estimate the angular velocity and linear acceleration with the update equations below.
\begin{equation}
\begin{split}
    \hat{\bm{\omega}}(t) = \bar{\textbf{y}}_\omega(t) - \hat{\bar{\textbf{b}}}_\omega(t) \\
    \hat{\textbf{a}}(t)  = \bar{\textbf{y}}_a(t) - \hat{\bar{\textbf{b}}}_a(t)
\end{split}
\end{equation}

@@

When selecting weight for gyroscope or for accelerometer when the selection of vehicle frame is flexible, we can minimize the output noise or bias drift rate. Below we take the bias drift as an example but similar idea can be applied to the output noise.

@@

Since most IMU have isotropic noise for all three axes, the combined variance simplifies to,
\begin{equation}\label{noise_reduction_asym}
    \bar{\sigma}^2 = \sum_i{a_i^2 \sigma_i^2}
\end{equation}

Further more, if using $n$ same IMU and equation weighting, the combined variance is simply,
\begin{equation}\label{noise_reduction_sym}
    \bar{\sigma}^2 = \sum_i{a_i^2 \sigma^2} = n \left(\frac{\sigma}{n}\right)^2 = \frac{\sigma^2}{n}
\end{equation}

\subsection{Selecting Weight for Vehicle Frame Placement}

Given that $\sum_i{a_i \textbf{r}_i} = \textbf{0}$, if there are only two IMUs, we can only place the vehicle frame on the line through the center of the two IMUs. If there are three IMUs, we can only place the vehicle frame on the plane formed by the three IMUs. If there are four non-coplanar IMUs, then in theory, we can place the vehicle frame anywhere. However, placing the vehicle frame outside of the convex hull of the IMUs will increase the output noise which is usually not desired.

When we have a reachable vehicle frame in mind, we need to solve for the weight. Assuming we have arbitrarily many IMUs, then the problem can be formulated as below.
\begin{equation}
\begin{split}
    \min_{w_i}{\frac{1}{2} \sum{w_i^2 \sigma_i^2}} \quad \text{s.t. }
    \begin{cases}
      \sum_i{w_i \textbf{r}_i} &= \textbf{0} \\
      \sum_i{w_i} &= 1
    \end{cases}
\end{split}
\end{equation}

Note that, in this section we will be using $w$ as weight instead of $a$ for convenience. Here $\textbf{r}_i$ is the relative position of the IMUs to the selected vehicle frame. Here we used the objective to reduce output noise as shown in section \ref{solve_weight_by_noise}. Since it is a quadratic programming problem, it is possible to find a close-form solution.

The Lagrangian of the problem is $\mathcal{L} = \frac{1}{2}\textbf{w}^T \bm{\Sigma} \textbf{w} + \bm{\lambda}^T \textbf{R}\textbf{w} + \beta \left( 1 - \textbf{1}^T \textbf{w} \right)$ = 0 where $\bm{\Sigma} = \text{diag}(\sigma_1^2, ..., \sigma_n^2)$, $\textbf{R} = \left[\begin{matrix} \textbf{r}_0 & ... & \textbf{r}_n\end{matrix}\right]$ and $\textbf{1}^T = \left[\begin{matrix}1 & ... & 1\end{matrix}\right]$. This gives the following KKT conditions.
\begin{equation}
\begin{cases}
  \bm{\Sigma}\textbf{w} + \textbf{R}^T \bm{\lambda} - \beta \textbf{1} &= \textbf{0} \\
  \textbf{R} \textbf{w} &= \textbf{0} \\
  \textbf{1}^T \textbf{w} &= 1
\end{cases}
\end{equation}

@@

Lastly since $\sum_i{w_i} = 1$ we need to normalize the intermediate solution to get the final weight for accelerometer.
\begin{equation}
    \textbf{w}^* = \frac{\hat{\textbf{w}}}{\textbf{1}^T \hat{\textbf{w}}}
\end{equation}

\section{Simulation}

\subsection{Data Generation}

First we test our finding in simulation. To generate the simulated data, we first use combination of sinusoidal functions to generate the ground truth angular velocity and linear acceleration in body frame as shown in Figure \ref{fig:omega} and Figure \ref{fig:accel}. Then we integrate it to generate the ground truth trajectory.
% as shown in Figure \ref{fig:trajectory}
For each IMU, we simulate its measurement given the relative transformation in body frame and added noises and biases. Each IMU is sampled at 100 Hz. We selected MPU6050, a common MEMS IMU in the mechatronics community as the example model. All of the above is done using utilities like \texttt{kinematicTrajectory} and \texttt{imuSensor} provided by MATLAB.

We also simulate observations of known landmark from a monocular camera. This turns the estimation problem into a localization problem such that the error is bounded. For each camera frame, we first sample random observations in imaging plane as well as depth (between 2 and 10 m). We then perform reprojection and transform it to inertial frame with the ground truth pose to generate the landmark position. Camera observations arrives at 2 Hz. We selected such a low update rate to allow error to propagate to better show the benefits of averaging.

% \begin{figure*}[ht]
%     \centering
%     \begin{subfigure}[b]{0.4\linewidth}
%         \centering
%         \includegraphics[width=5.8cm]{figures/simulated_omega.png}
%         \caption{Simulated Angular Velocity}
%         \label{fig:omega}
%     \end{subfigure}
%     \begin{subfigure}[b]{0.4\linewidth}
%         \centering
%         \includegraphics[width=5.8cm]{figures/simulated_accel.png}
%         \caption{Simulated Linear Acceleration}
%         \label{fig:accel}
%     \end{subfigure}
%     \begin{subfigure}[b]{0.4\linewidth}
%         \centering
%         \includegraphics[width=5.8cm]{figures/simulated_trajectory.png}
%         \caption{Ground truth Trajectory from Integration}
%         \label{fig:trajectory}
%     \end{subfigure}
% \caption{Simulated Data Generation with MATLAB}
% \label{fig:switching_diagram}
% \end{figure*}

\begin{figure}
    \centering
    \includegraphics[width=0.85\linewidth]{figures/simulated_omega.png}
    \caption{Simulated Angular Velocity}
    \label{fig:omega}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=0.85\linewidth]{figures/simulated_accel.png}
    \caption{Simulated Linear Acceleration}
    \label{fig:accel}
\end{figure}
% \begin{figure}
%     \centering
%     \includegraphics[width=0.85\linewidth]{figures/simulated_trajectory.png}
%     \caption{Ground truth Trajectory from Integration}
%     \label{fig:trajectory}
% \end{figure}

\subsection{Implementation}

For estimator we implemented the \textit{left-}Invariant Extended Kalman Filter (IEKF) \cite{IEKF} as it is more accurate compared to standard EKF methods. At initialization stage, given all IMU poses, we compute the weighting that places the VIMU frame at body frame. Using the weights, we also calculate the combined noises and bias drift rates. When new IMU measurements arrive, they are weighted averaged and then feed into the estimator same as single IMU case. When camera observations arrives, the 2D observations and 3D landmark positions are used for correction step.

\subsection{Experiment Setup}

For performance, we compared results of multiple different IMU configurations summarized in Table \ref{tab:imu_config}. Here symmetric configurations are placing IMUs along the principle axes with an offset of 1 m and random rotation. Assymmetric configurations use IMUs perturbed from symmetric configuration such that different weighting than 1 is required to have VIMU frame at origin of body frame. We place the VIMU frame same as body frame for all configurations for fair comparison of performance. Pose of all IMUs used are summarized in Table \ref{tab:imu_pose}.

\begin{table}[h!]
\centering
\caption{Different IMU Configurations}
\label{tab:imu_config}
\begin{tabular}{cccc}
\toprule
\textbf{Configuration} & \textbf{Category} & \textbf{\# of IMUs} & \textbf{IMU used}  \\
\midrule
IMU-S0 & Baseline  & 1 & 0 \\
\midrule
IMU-S2 & Symmetric & 2 & 1, 2 \\
IMU-S4 & Symmetric & 4 & 1, 2, 3, 4 \\
IMU-S6 & Symmetric & 6 & 1, 2, 3, 4, 5, 6 \\
\midrule
IMU-A2 & Asymmetric & 2 & 11, 12 \\
IMU-A4 & Asymmetric & 4 & 11, 12, 13, 14 \\
IMU-S6 & Asymmetric & 6 & 11, 12, 13, 14, 15, 16 \\

\bottomrule
\end{tabular}
\end{table}

\begin{table}[h!]
\centering
\caption{Pose of IMUs}
\label{tab:imu_pose}
\begin{tabular}{ccc}
\toprule
\textbf{IMU ID} & \textbf{orientation} & \textbf{position} \\
\midrule
0      & $1.000 +0.000i +0.000j + 0.000k$ & [$+0.0, +0.0, +0.0$] \\
\midrule
1      & $0.633 +0.065i -0.025j +0.771k$ & [$+1.0, +0.0, +0.0$] \\
2      & $0.395 +0.795i +0.322j +0.328k$ & [$-1.0, +0.0, +0.0$] \\
3      & $0.533 -0.314i +0.671j +0.409k$ & [$+0.0, +1.0, +0.0$] \\
4      & $0.600 +0.323i +0.620j +0.390k$ & [$+0.0, -1.0, +0.0$] \\
5      & $0.231 +0.495i +0.343j -0.764k$ & [$+0.0, +0.0, +1.0$] \\
6      & $0.744 -0.473i +0.355j +0.310k$ & [$+0.0, +0.0, -1.0$] \\
\midrule
11     & $0.045 -0.112i +0.670j +0.733k$ & [$+0.9, +0.0, +0.0$] \\
12     & $0.223 -0.059i +0.929j -0.289k$ & [$-1.2, +0.0, +0.0$] \\
13     & $0.119 +0.962i +0.245j -0.031k$ & [$+0.3, +1.1, +0.0$] \\
14     & $0.605 +0.708i +0.335j +0.141k$ & [$+0.2, -0.7, +0.0$] \\
15     & $0.004 -0.299i +0.894j +0.333k$ & [$-0.1, +0.2, +1.3$] \\
16     & $0.598 -0.094i -0.669j -0.432k$ & [$+0.4, -0.3, -1.1$] \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Results}

The results are summarized in Table \ref{tab:sim_result}. As more and more IMUs are fused, the mean errors decrease. This matches with the expectation as the combined noises from fused IMU are attenuated as shown by equation \ref{noise_reduction_asym} and \ref{noise_reduction_sym}. Comparing symmetric and asymmetric configuration, the error is marginally larger for asymmetric configuration. This can be explained by the non-uniform weighting which gives higher level of noises than using equal weighting. Overall, the result suggests averaging between IMUs works for both gyroscope and accelerometer measurements disregarding the orientation and separation and the combined measurements are more accurate.

\begin{table}[h!]
\centering
\caption{Simulation Results}
\label{tab:sim_result}
\begin{tabular}{ccccccc}
\toprule
\textbf{Configuration} & \multicolumn{2}{c}{\textbf{Attitude Err (rad)}} & \multicolumn{2}{c}{\textbf{Position Err (m)}} \\
\cmidrule(lr){2-3} \cmidrule(lr){4-5}
& MAE & RMSE & MAE & RMSE \\
\midrule
IMU-S0 & 0.01568 & 0.01873 & 0.11189 & 0.13260 \\
\midrule
IMU-S2 & 0.01278 & 0.01486 & 0.10029 & 0.11540 \\
IMU-S4 & 0.01122 & 0.01319 & 0.09223 & 0.10598 \\
IMU-S6 & 0.01028 & 0.01176 & 0.08950  & 0.10271 \\
\midrule
IMU-A2 & 0.01306 & 0.01514 & 0.10055 & 0.11685 \\
IMU-A4 & 0.01129 & 0.01285 & 0.09222 & 0.10554 \\
IMU-A6 & 0.01083 & 0.01221 & 0.08944 & 0.10077 \\
\bottomrule
\end{tabular}
\end{table}

\section{Real-World Experiment}

For real-world experiment, we choose to perform visual-inertial odometry to understand the performance of averaging in real-life scenario.

@@

Most public datasets only contain single IMU. The dataset we find that contains more than one IMU is the PennCOSYVIO dataset \cite{penncosyvio}. This dataset contains three IMUs and multiple cameras. For our experiment, we choose to only use the VI sensor camera as they were provided with complete and correct information regarding intrinsic, extrinsic, and distortion information. We also only used the left camera of the VI sensor such that it forms the minimal configuration with single IMU. This is to not hide the impact of IMUs on performance. We select the VI sensor IMU as the main IMU as it is more accurate compared to the IMUs in the two Tango devices. We will be comparing the performance of using only the VI sensor IMU vs VIMU from averaging all three IMUs.

\subsection{Implementation}

For VIO implementation, we choose to use OpenVINS \cite{openvins} to test the different IMU configurations. OpenVINS is very powerful as it can also estimate camera and IMU intrinsic as well as time offset and extrinsic between IMU and camera. After some tuning, we found that enabling extrinsic and time offset estimation gives the best result over many repeated runs. Also due to the PennCOSYVIO dataset doesn't have a initialization phase when the vehicle is at rest, as a result, we enabled dynamic initialization in OpenVINS.

\subsection{Setup}

To demonstrate the the weighting works as expected, we designed four configuration to test. 1) Single IMU with weight $[\begin{matrix} 1.0 & 0.0 & 0.0 \end{matrix}]$. Here the first weight is for VI sensor IMU, second weight for top Tango IMU, and the third weight is for bottom Tango IMU. 2) Averaged VIMU with weight $[\begin{matrix} 0.33 & 0.33 & 0.33 \end{matrix}]$. 3) VIMU closest to body frame origin with weight $[\begin{matrix} 0.4944 & 0.1546 & 0.3509 \end{matrix}]$. This places the VIMU at $[\begin{matrix} 0.0186 & -0.0012 & -0.0046 \end{matrix}]$ in body frame. 4) We also designed a case when the weight goes negative $[\begin{matrix} 1.2 & -0.1 & -0.1 \end{matrix}]$. This places the VIMU outside of the triangle formed by the three IMUs.

\subsection{Results}

The results are summarized in Table \ref{tab:vio_result}. Example trajectories against ground truth are plotted in \textcolor{red}{Figure} \ref{}.

From the result, different configurations perform similarly. Error reduction from averaging as in simulation is not seen. This is believed to be due to various imperfections from the real-world dataset.

1) Poor calibration. According to the paper of the PennCOSYVIO dataset, the extrinsic calibration is only done between the cameras (for VI sensor only the left camera is calibrated). The IMU positions are assumed from the factory calibration relative to their respective cameras. The chain of transformation can introduces calibration error when averaging the two Tango IMUs with the VI sensor IMU. The VIMU is expected to have less accurate calibration when used with the VI sensor compared to the VI sensor IMU.

2) No static data. Another issue with the dataset is that there is no static stage at the beginning. The initial static stage is crucial for accurate VIO as it allows estimating the initial biases as well as provide known zero initial velocity. As a result, in OpenVINS we have to rely on dynamic initialization which is less accurate compared to static initialization.

3) Vibration and flexing in the fixture. Due to the large separation between all sensors, the physical vibration and deflection in the system can introduce noise and error as well.

It is expected that the noise improvement from averaging is overwhelmed by the imperfection mentioned above. However, from the result we can still conclude that averaging across multiple IMUs with separation does work as expected giving the freedom of placing VIMU frame at preferred location which may previously be not accessible.
